### 2012-09-17 doc/prelim/document/chapters/linear_problem/linear_problem.tex
prelim updates

\section{Preliminaries}
\label{sec:linear_preliminaries}
We seek solutions of the general linear problem in the following form:
\begin{equation}
  \ve{A} \ve{x} = \ve{b}\:,
  \label{eq:linear_problem}
\end{equation}
where $\ve{A} \in \mathbb{R}^{N \times N}$ is a matrix operator such
that $\ve{A} : \mathbb{R}^{N} \rightarrow \mathbb{R}^{N}$, $\ve{x} \in
\mathbb{R}^N$ is the solution vector, and $\ve{b} \in \mathbb{R}^N$ is
the forcing term. The solutions to Eq~(\ref{eq:linear_problem}) will
be generated by inverting $\ve{A}$ either directly or indirectly:
\begin{equation}
  \ve{x} = \ve{A}^{-1} \ve{b}
  \label{eq:linear_problem_solution}\:.
\end{equation}
In addition we can define the residual:
\begin{equation}
  \ve{r} = \ve{b} - \ve{A}\ve{x}\:,
  \label{eq:linear_residual}
\end{equation}
such that an exact solution to $\ve{x}$ has been found when
$\ve{r}=\ve{0}$.  From the statement in
Eq~(\ref{eq:linear_problem_solution}) we can already place a
restriction on $\ve{A}$ by requiring that it be \textit{nonsingular},
meaning that we can in fact compute $\ve{A}^{-1}$. In this work we
will focus our efforts on approximately inverting the operator through
various means.

@@

General to both matrices and vectors, \textit{norms} are a mechanism
for collapsing objects of many elements to a single value. Per
LeVeque's text \citep{leveque_2007}, the q-norm of a vector is defined
as:
\begin{equation}
  ||\ve{v}||_q = \Bigg[ \sum_{i=1}^N |v_i|^q \Bigg]^{1/q},\ \ve{v} \in
  \mathbb{R}^N\:,\ q \in \mathbb{Z}^+
  \label{eq:q_norm}
\end{equation}
where ${v_i}$ is the $i^{th}$ component of the vector. Depending on
the value chosen for $q$, local or global qualities of the vector may
be obtained. For example, $q=2$ provides the root of a quadrature sum
of all elements in the vector giving a global measure of the vector
while $q=\infty$ gives the maximum value in the vector, a local
quantity that does not give information regarding the other elements
in the vector.

We can also compute the norm of a matrix by inferring from the norm of
the vector on which it is operating. Per LeVeque, we search for a
constant that is equivalent to $||\ve{A}||$:
\begin{equation}
  ||\ve{A}\ve{x}|| \leq C ||\ve{x}||\:,
  \label{eq:matrix_norm_inequality}
\end{equation}
where the minimum value of $C$ that satisfies
Eq~(\ref{eq:matrix_norm_inequality}) is equivalent to $||\ve{A}||$
and is valid $\forall \ve{x} \in \mathbb{R}^N$. The general
definition in Eq~(\ref{eq:matrix_norm_inequality}) can be expanded in
simple terms for common norms including the infinity norm:
\begin{equation}
  ||\ve{A}||_{\infty} = \max_{1 \leq i \leq N} \sum^N_{j=1}|a_{ij}|\:,
  \label{eq:matrix_infinity_norm}
\end{equation}
and the 2-norm:
\begin{equation}
  ||\ve{A}||_{2} = \sqrt{\rho(\ve{A}^T\ve{A})}\:,
  \label{eq:matrix_2_norm}
\end{equation}
where $\rho$ is the spectral radius as defined in
Eq~(\ref{eq:spectral_radius}).

@@

\subsubsection{Preconditioning Stochastic Methods}
\label{subsubsec:stochastic_preconditioning}
In most cases, at least a minimal amount of \textit{preconditioning}
of the linear system will be required in order to use the class of
stochastic methods described. Although these methods have no symmetry
requirements for convergence, they do require that the spectral radius
of the iteration matrix be less than one. To achieve this, a Jacobi
precondtioner, a form of left preconditioning, is used such that the
preconditioning matrix $\ve{D}$ is:
\begin{equation}
  \ve{M} = diag(\ve{A})\:,
  \label{eq:jacobi_preconditioner}
\end{equation}
such that its application means we are instead solving the following
linear system:
\begin{equation}
  \ve{M}^{-1}\ve{A}\ve{x} = \ve{M}^{-1}\ve{b}\:.
  \label{eq:jacobi_precond_linear_problem}
\end{equation}
Next, we can apply MCSA to solve
Eq~(\ref{eq:jacobi_precond_linear_problem}): 
\begin{subequations}
  \begin{gather}
    \ve{x}^{k+1/2} = (\ve{I} - \ve{M}^{-1}\ve{A})\ve{x}^k +
    \ve{M}^{-1}\ve{b}\:,\\ \ve{r}^{k+1/2} = \ve{M}^{-1}\ve{b} -
    \ve{M}^{-1}\ve{A}\ve{x}^{k+1/2}\:,\\ \ve{M}^{-1}\ve{A}\delta\ve{x}^{k+1/2}
    = \ve{r}^{k+1/2}\:,\\ \ve{x}^{k+1} = \ve{x}^{k+1/2} + \delta
    \ve{x}^{k+1/2}\:.
    \label{eq:jacobi_preconditioned_mcsa}
  \end{gather}
\end{subequations}
Choosing Jacobi preconditioning with MCSA is advantageous for several
reasons. First, $\rho(\ve{I} - \ve{M}^{-1}\ve{A}) < 1$ is true for
all $\ve{A}$ and is easy to formulate as the inversion of $\ve{M}$ is
trivial. Second, because the adjoint Monte Carlo method used within MCSA to
compute the correction operates on a linear problem with the
preconditioned operator, then $\ve{H}$ in the adjoint solver will have
a zero term in each of its diagonal elements, thereby eliminating
all in-state transitions during the random walk sequence. Because of
this, Jacobi preconditioning should always be performed, regardless of
any other preconditioning that is applied to the system.

>>>>

\section{Preliminaries}
\label{sec:linear_preliminaries}
We seek solutions of the general linear problem in the following form:
\begin{equation}
  \ve{A} \ve{x} = \ve{b}\:,
  \label{eq:linear_problem}
\end{equation}
where $\ve{A} \in \mathbb{C}^{N \times N}$ is a matrix operator such
that $\ve{A} : \mathbb{C}^{N} \rightarrow \mathbb{C}^{N}$, $\ve{x} \in
\mathbb{C}^N$ is the solution vector, and $\ve{b} \in \mathbb{C}^N$ is
the forcing term. The solutions to Eq~(\ref{eq:linear_problem}) will
be generated by inverting $\ve{A}$ either directly or indirectly:
\begin{equation}
  \ve{x} = \ve{A}^{-1} \ve{b}
  \label{eq:linear_problem_solution}\:.
\end{equation}
In addition we can define the residual:
\begin{equation}
  \ve{r} = \ve{b} - \ve{A}\ve{x}\:,
  \label{eq:linear_residual}
\end{equation}
such that an exact solution to $\ve{x}$ has been found when
$\ve{r}=\ve{0}$.  From the statement in
Eq~(\ref{eq:linear_problem_solution}) we can already place a
restriction on $\ve{A}$ by requiring that it be \textit{nonsingular},
meaning that we can in fact compute $\ve{A}^{-1}$. In this work we
will focus our efforts on approximately inverting the operator through
various means.

@@

General to both matrices and vectors, \textit{norms} are a mechanism
for collapsing objects of many elements to a single value. Per
LeVeque's text \citep{leveque_2007}, the q-norm of a vector is defined
as:
\begin{equation}
  ||\ve{v}||_q = \Bigg[ \sum_{i=1}^N |v_i|^q \Bigg]^{1/q},\ \ve{v} \in
  \mathbb{C}^N\:,\ q \in \mathbb{Z}^+
  \label{eq:q_norm}
\end{equation}
where ${v_i}$ is the $i^{th}$ component of the vector. Depending on
the value chosen for $q$, local or global qualities of the vector may
be obtained. For example, $q=2$ provides the root of a quadrature sum
of all elements in the vector giving a global measure of the vector
while $q=\infty$ gives the maximum value in the vector, a local
quantity that does not give information regarding the other elements
in the vector.

We can also compute the norm of a matrix by inferring from the norm of
the vector on which it is operating. Per LeVeque, we search for a
constant that is equivalent to $||\ve{A}||$:
\begin{equation}
  ||\ve{A}\ve{x}|| \leq C ||\ve{x}||\:,
  \label{eq:matrix_norm_inequality}
\end{equation}
where the minimum value of $C$ that satisfies
Eq~(\ref{eq:matrix_norm_inequality}) is equivalent to $||\ve{A}||$
and is valid $\forall \ve{x} \in \mathbb{C}^N$. The general
definition in Eq~(\ref{eq:matrix_norm_inequality}) can be expanded in
simple terms for common norms including the infinity norm:
\begin{equation}
  ||\ve{A}||_{\infty} = \max_{1 \leq i \leq N} \sum^N_{j=1}|a_{ij}|\:,
  \label{eq:matrix_infinity_norm}
\end{equation}
and the 2-norm:
\begin{equation}
  ||\ve{A}||_{2} = \sqrt{\rho(\ve{A}^T\ve{A})}\:,
  \label{eq:matrix_2_norm}
\end{equation}
where $\rho$ is the spectral radius as defined in
Eq~(\ref{eq:spectral_radius}).

@@

\subsubsection{Preconditioning Stochastic Methods}
\label{subsubsec:stochastic_preconditioning}
In most cases, at least a minimal amount of \textit{preconditioning}
of the linear system will be required in order to use the class of
stochastic methods described. Although these methods have no symmetry
requirements for convergence, they do require that the spectral radius
of the iteration matrix be less than one. To achieve this, a Jacobi
precondtioner, a form of left preconditioning, is used such that the
preconditioning matrix $\ve{M}$ is:
\begin{equation}
  \ve{M} = diag(\ve{A})\:,
  \label{eq:jacobi_preconditioner}
\end{equation}
such that its application means we are instead solving the following
linear system:
\begin{equation}
  \ve{M}^{-1}\ve{A}\ve{x} = \ve{M}^{-1}\ve{b}\:.
  \label{eq:jacobi_precond_linear_problem}
\end{equation}
Next, we can apply MCSA to solve
Eq~(\ref{eq:jacobi_precond_linear_problem}): 
\begin{subequations}
  \begin{gather}
    \ve{x}^{k+1/2} = (\ve{I} - \ve{M}^{-1}\ve{A})\ve{x}^k +
    \ve{M}^{-1}\ve{b}\:,\\ \ve{r}^{k+1/2} = \ve{M}^{-1}\ve{b} -
    \ve{M}^{-1}\ve{A}\ve{x}^{k+1/2}\:,\\ \ve{M}^{-1}\ve{A}\delta\ve{x}^{k+1/2}
    = \ve{r}^{k+1/2}\:,\\ \ve{x}^{k+1} = \ve{x}^{k+1/2} + \delta
    \ve{x}^{k+1/2}\:.
    \label{eq:jacobi_preconditioned_mcsa}
  \end{gather}
\end{subequations}
Choosing Jacobi preconditioning with MCSA is advantageous for several
reasons. First, $\rho(\ve{I} - \ve{M}^{-1}\ve{A}) < 1$ is true for
all $\ve{A}$ and is easy to formulate as the inversion of $\ve{M}$ is
trivial. Second, because the adjoint Monte Carlo method used within MCSA to
compute the correction operates on a linear problem with the
preconditioned operator, then $\ve{H}$ in the adjoint solver will have
a zero term in each of its diagonal elements, thereby eliminating
all in-state transitions during the random walk sequence. Because of
this, Jacobi preconditioning should always be performed, regardless of
any other preconditioning that is applied to the system.

### 2012-09-17 doc/prelim/document/chapters/parallel_mcsa/parallel_mcsa.tex
updates to linear systems chapter
updates to prelim, projection methods section is complete
prelim updates

\chapter{Parallelization of the MCSA Algorithm}
\label{ch:parallel_mcsa}

@@

\section{Current Parallel Methodologies for Solving Linear Systems}
\label{sec:current_parallel_methods}

>>>>

\chapter{Parallelization of the MCSA Algorithm for Linear Problems}
\label{ch:parallel_mcsa}

@@

\section{Current Parallel Methodologies for Solving Linear Systems}
\label{sec:current_parallel_methods}
Modern parallel implementations of linear solver methods rely heavily
on capabilities provided by general linear algebra frameworks. In
particular, Krylov methods that are formulated through matrix-vector
multiply operations, are easily formulated in parallel using these
frameworks.
